Explore and Summarize Prosper Loan Data by Michael Erickson
========================================================
Prosper is a peer-to-peer lending company, where individuals can invest in
personal loans and request to borrow money 
\([Wikipedia](https://en.wikipedia.org/wiki/Prosper_Marketplace)\). The data set
contains nearly 114,000 observations of loans with extensive information about
each loan, including payment information.

```{r Load Packages, echo=FALSE, message=FALSE, warning=FALSE}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(gridExtra)
library(maps)
library(viridis)
```

```{r Load Data, echo=FALSE}
# Load the Data
loans <- read.csv('prosperLoanData.csv')
```

#Data Transformation
Twenty-one values for ListingCategory are stored as integers 0 to 20. These are 
converted to Factors with labels indicating their use.
```{r Listing Category Factor, echo=TRUE}
# Transform ListingCategory to a Factor
loans$ListingCategory <- factor(loans$ListingCategory..numeric.,
                                labels = c("Not Available",
  "Debt Consolidation", "Home Improvement", "Business", "Personal Loan",
  "Student Use", "Auto", "Other", "Baby&Adoption", "Boat", "Cosmetic Procedure",
  "Engagement Ring", "Green Loans", "Household Expenses", "Large Purchases",
  "Medical/Dental", "Motorcycle", "RV", "Taxes", "Vacation", "Wedding Loans"))
```

To facilitate plotting and filtering outlying Loan Amounts, a boolean is added
indicating if the Amount is greater than third quartile plus 1.5 times the
interquartile range (IQR) or lower than first quartile minus 1.5 times the IQR.
```{r Transform Data, echo=TRUE}
#Add Outlier for High loan amounts
loans$LoanOriginalAmount.outlier.high <-
  (loans$LoanOriginalAmount > quantile(loans$LoanOriginalAmount, .75)
    + 1.50*IQR(loans$LoanOriginalAmount))

#Add Outlier for Low loan amounts
loans$LoanOriginalAmount.outlier.low <-
  (loans$LoanOriginalAmount < quantile(loans$LoanOriginalAmount, .25)
    - 1.50*IQR(loans$LoanOriginalAmount))
```

# Univariate Plots
### Loan Status
```{r Loan Status, echo=FALSE}

# Plot Listings by Loan Status
ggplot(loans, aes(x=loans$LoanStatus)) +
  geom_bar() +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.4)) +
  labs(title='Loan Status',
       x=NULL,
       y='Number of Listings')
```

Most loans appear to be current in their payments or to sucessfully complete. 
It would be interesting to explore correlation between Chargedoff / Defaulted 
loans and ProsperRating or ProsperScore (the Prosper evaluation of a borrower's 
credit worthiness).

### Listing Category
```{r Listing Category, echo=FALSE}
#Plot percent of listings by Category
ggplot(loans, aes(x=loans$ListingCategory, y=..count../n, n=nrow(loans))) +
  geom_bar() +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.4)) +
  labs(title='Listing Category',
       x=NULL,
       y='Percent of Listings')
```

Debt consolidation is the predominant reason on Prosper for borrowing money. 
Over half of all listings are for Debt Consolidation. I would like to explore 
listing category against borrower income to determine if different levels of 
income tend to borrow for different reasons.

### Income Range
```{r Income Range, echo=FALSE}
#Plot percent of listings by income level
ggplot(loans, aes(x=loans$IncomeRange, y=..count../n, n=nrow(loans))) +
  geom_bar() +
  scale_y_continuous(labels = scales::percent) +
  scale_x_discrete(limits=c('Not displayed', 'Not employed', '$0', '$1-24,999',
                            '$25,000-49,999', '$50,000-74,999',
                            '$75,000-99,999', '$100,000+')) +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.4)) +
  labs(title='Income Levels of Listings',
       x=NULL,
       y='Percent of Listings')
```

Prosper borrowers appear to be significantly more affluent than the general 
population of the U.S.
|     Bracket    | Prosper | United States |
|:--------------:|:-------:|:-------------:|
| Less than $25K |      6% |           35% |
| $25K to $50K   |     28% |           36% |
| $50K to $75K   |     27% |           16% |
| $75K to $100K  |     15% |            6% |
| Over $100K     |     15% |            7% |
(Prosper values exclude borrowers with income not displayed, who are unemployed, 
or who have no income. United States values are for people aged 25-64 with income.)

Prosper borrowers could be older and likely to earn more. Prosper borrowers are 
certainly self-selecting as the type of person who would take a loan out through 
an online lender. Prosper's lending requirements could also bias to more affluent 
borrowers.

Reference: [Personal Income in the United States](https://en.wikipedia.org/wiki/Personal_income_in_the_United_States#Income_distribution)

### Loan Amount
```{r Loan Amount, echo=FALSE}

#boxplot of LoanOriginalamount
n1 <- 
  ggplot(loans, aes(x='', y=LoanOriginalAmount)) +
    geom_boxplot(outlier.shape=NA) +
    geom_jitter(data=dplyr::filter(loans,
      LoanOriginalAmount.outlier.high == TRUE 
      | LoanOriginalAmount.outlier.low == TRUE),
      color='red', width=0.1, alpha=0.1) +
    labs(title='Box Plot of LoanOriginalAmount',
         x=NULL,
         y='LoanOriginalAmount')

#histogram of LoanOriginalamount
n2 <-
  qplot(loans$LoanOriginalAmount,
        geom='histogram',
        binwidth=1000,
        col=I('white'),
        main='Histogram of LoanOriginalAmount',
        xlab='Loan Amount')  

#tabular data of loan amount frequency divisible by 10s, 100s, 1000s, 10000s
n3 <- tibble('Divisor'=integer(),
             'Distinct_Values'=integer(),
             'Number_of_Loans'=integer())

for (i in 1:4) { #powers of 10 math and summary
  loans %>%
  filter(LoanOriginalAmount%%(10^i) == 0) %>%
  summarize(sum=sum(n()), count=n_distinct(LoanOriginalAmount)) %>%
  {add_row(n3, 'Divisor' = (10^i),
                          'Distinct_Values'= .$count,
                          'Number_of_Loans' = .$sum)} -> n3
}


#histogram of last digit of loan amount
n4 <-
  qplot(loans$LoanOriginalAmount%%10,
      geom='histogram',
      binwidth=1,
      col=I('white'),
      main='Histogram of Last Digit',
      xlab='Last Digit')


#place in grid
grid.arrange(n1, n2, tableGrob(n3), n4, ncol=2)


#cleanup
rm(n1, n2, n3, i, n4)
```

#### Loan Amount Summary Statistics
```{r Loan Amount 2, echo=FALSE}
summary(loans$LoanOriginalAmount)
```

```{r Loan Amount 3, echo=FALSE}

loans %>%
  count(LoanOriginalAmount) %>%
  arrange(desc(n)) %>%
  mutate(pct=n/sum(n)) %>%
  head(10)
```

Prosper currently offers loans between $2,000 and $40,000. This data set appears 
to have data before that change was implemented, as loans in the set are bounded 
$1,000-$35,000. There is bias towards "large, round" loan amounts, indicating 
borrowers tend to select an amount near what they need rather than precisely. 
About 74% (84,000 loans of 114,000 total loans) of loans are a multiple of 
thousands.

$4,000 is the most commonly borrowed amount. I'm curious if the amount is also 
associated with a specific type of purchase (other than debt consolidation). 
$15,000 and $10,000 are also frequently occuring amounts. These top three account 
for one-third of all loans, and the top 10 loan amounts account for 61%.

Loan amounts skew right, which is typical of a personal income distribution, and 
I would expect loan amount to have a positive correlation with borrower income.

### Loan Amounts by State
```{r Loan Amount by State, echo=FALSE}

#summarize average loan amount by state
avg_loan_by_state <-
loans %>%
  group_by(BorrowerState) %>%
  summarize(n=n(), loan_amount=sum(LoanOriginalAmount)) %>%
  mutate(state_name = tolower(state.name[match(BorrowerState, state.abb)]),
         amount = loan_amount/n)

#load us state map data
us <- map_data('state')

#plot us map with average loan amounts
ggplot() +
  geom_blank(data=us, aes(long, lat)) +
  geom_map(data=us, map=us, aes(map_id=region),
           fill='black', color='black', size=0.01) +
  geom_map(data=avg_loan_by_state, map=us,
           aes(fill=amount, map_id=state_name), 
           color='grey40', size=0.01) +
  coord_map('polyconic') +
  labs(x=NULL, y=NULL, title='Average Loan Amount by State') +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks   = element_blank(),
        axis.text    = element_blank()) +
  scale_fill_viridis(name='Amount', option='inferno',
                     label=scales::dollar)

#cleanup
rm(us, avg_loan_by_state)
```

I would expect that average loan amount and median income by state would be
positively correlated. This is the case in some states, such as California,
Texas,and much of the northeast, but does not hold true in many other areas.

For instance, North Dakota has a relatively high median income, but
comparatively low average loan value. Conversely, Louisiana has median income
comparable to other gulf states, but higher average loan values.

This comparison is not rigorous, and I don't expect to dig in to it further in 
this analysis, but it is an interesting map.

Reference: [Census: Median household income, 2012-2016](https://www.census.gov/quickfacts/fact/map/US/INC110216)

# Univariate Analysis
For each plot above, I provided interesting findings about variables I explored. 
Here, I discuss what I initially intend to explore further:
- relationship between Prosper rating / score of creditworthiness and the \
default rate 
- reason for borrowing money against borrower income level; are reasons \
consistent or varied?
- loan amount vs. income

The Prosper data set is well-structured with loan observations on each row and 
many interesting variables in columns. There is enough information to determine 
if data Prosper collects is sufficient to determine credit risk and how
predictive Prosper's model is of repayment.

I performed slight cleanup on the data to:
- convert loan listing categories from numeric to string factors
- add informational boolean values to high and low loan value outliers

Other than the above, I performed no adjustment to underlying data values.


# Bivariate Plots

> **Tip**: Based on what you saw in the univariate plots, what relationships
between variables might be interesting to look at in this section? Don't limit
yourself to relationships between a main output feature and one of the
supporting variables. Try to look at relationships between supporting variables
as well.

```{r echo=FALSE, Bivariate_Plots}

```

# Bivariate Analysis

> **Tip**: As before, summarize what you found in your bivariate explorations
here. Use the questions below to guide your discussion.

### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?

### What was the strongest relationship you found?


# Multivariate Plots Section

> **Tip**: Now it's time to put everything together. Based on what you found in
the bivariate plots section, create a few multivariate plots to investigate
more complex interactions between variables. Make sure that the plots that you
create here are justified by the plots you explored in the previous section. If
you plan on creating any mathematical models, this is the section where you
will do that.

```{r echo=FALSE, Multivariate_Plots}

```

# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

### Were there any interesting or surprising interactions between features?

### OPTIONAL: Did you create any models with your dataset? Discuss the \
strengths and limitations of your model.

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, Plot_One}

```

### Description One


### Plot Two
```{r echo=FALSE, Plot_Two}

```

### Description Two


### Plot Three
```{r echo=FALSE, Plot_Three}

```

### Description Three

------

# Reflection

> **Tip**: Here's the final step! Reflect on the exploration you performed and
the insights you found. What were some of the struggles that you went through?
What went well? What was surprising? Make sure you include an insight into
future work that could be done with the dataset.

> **Tip**: Don't forget to remove this, and the other **Tip** sections before
saving your final work and knitting the final report!