Explore and Summarize Prosper Loan Data by Michael Erickson
========================================================
Prosper is a peer-to-peer lending company, where individuals can invest in
personal loans and request to borrow money 
([Wikipedia](https://en.wikipedia.org/wiki/Prosper%5FMarketplace)). The data set
contains nearly 114,000 observations of loans with extensive information about
each loan, including payment information.

```{r Load Packages, echo=FALSE, message=FALSE, warning=FALSE}
# Load all of the packages that you end up using in your analysis in this code
# chunk.

# Notice that the parameter "echo" was set to FALSE for this code chunk. This
# prevents the code from displaying in the knitted HTML output. You should set
# echo=FALSE for all code chunks in your file, unless it makes sense for your
# report to show the code that generated a particular plot.

# The other parameters for "message" and "warning" should also be set to FALSE
# for other code chunks once you have verified that each plot comes out as you
# want it to. This will clean up the flow of your report.

library(ggplot2)
library(dplyr)
library(gridExtra)
library(maps)
library(viridis)
library(GGally)
library(knitr)
```

```{r Load Data, echo=FALSE}
# Load the Data
loans <- read.csv('prosperLoanData.csv')
```

#Data Transformation
Remove extraneous characters from variable names.
```{r Rename Variables, echo=TRUE}
# remove extraneous characters from variable names
loans <- rename(loans,
                'ProsperRatingAlpha' = 'ProsperRating..Alpha.',
                'ProsperRatingNumeric' = 'ProsperRating..numeric.')
```

Twenty-one values for ListingCategory are stored as integers 0 to 20. These are 
converted to Factors with labels indicating their use. Loan Terms are present in 
three discrete values (12, 36, 60), so I will convert them as well.
```{r Transform to Factor, echo=TRUE}
# Transform ListingCategory to a Factor
loans$ListingCategory <- factor(loans$ListingCategory..numeric.,
                                labels = c("Not Available",
  "Debt Consolidation", "Home Improvement", "Business", "Personal Loan",
  "Student Use", "Auto", "Other", "Baby&Adoption", "Boat", "Cosmetic Procedure",
  "Engagement Ring", "Green Loans", "Household Expenses", "Large Purchases",
  "Medical/Dental", "Motorcycle", "RV", "Taxes", "Vacation", "Wedding Loans"))

#Loan Terms to factors
loans$Term.f <- factor(loans$Term)
```

To facilitate plotting and filtering outlying Loan Amounts, a boolean is added
indicating if the Amount is greater than third quartile plus 1.5 times the
interquartile range (IQR) or lower than first quartile minus 1.5 times the IQR.
```{r Outliers, echo=TRUE}
# Add Outlier for High loan amounts
loans$LoanOriginalAmount.outlier.high <-
  (loans$LoanOriginalAmount > quantile(loans$LoanOriginalAmount, .75)
    + 1.50*IQR(loans$LoanOriginalAmount))

# Add Outlier for Low loan amounts
loans$LoanOriginalAmount.outlier.low <-
  (loans$LoanOriginalAmount < quantile(loans$LoanOriginalAmount, .25)
    - 1.50*IQR(loans$LoanOriginalAmount))
```

Handle ListingCreationDate as a POSIX datetime to simplify calculations.
```{r Listing Creation Date, echo=TRUE}
loans$ListingCreationDate.dt <- as.POSIXct(loans$ListingCreationDate,
                                           format='%Y-%m-%d %H:%M:%S')

june2009 <- as.POSIXct('2009-07-01', tz='UTC', format='%Y-%m-%d')
```

Order Factors logically for display in plots.
```{r Order Factors, echo=TRUE}
# Order ProsperRating high to low
loans$ProsperRatingAlpha <-
  factor(loans$ProsperRatingAlpha,
         levels=c('AA', 'A', 'B', 'C', 'D', 'E', 'HR', 'NC', ''))

# Order IncomeRange low to high
loans$IncomeRange <-
  factor(loans$IncomeRange,
       levels=c('Not displayed', 'Not employed', '$0', '$1-24,999',
                        '$25,000-49,999', '$50,000-74,999',
                        '$75,000-99,999', '$100,000+'))

# Order LoanStatus by groups
loans$LoanStatus <-
  factor(loans$LoanStatus,
         levels=c('Cancelled', 'Completed', 'Current', 'FinalPaymentInProgress',
                  'Chargedoff', 'Defaulted', 'Past Due (1-15 days)',
                  'Past Due (16-30 days)', 'Past Due (31-60 days)',
                  'Past Due (61-90 days)', 'Past Due (91-120 days)',
                  'Past Due (>120 days)'))
```

# Univariate Plots
### Loan Status
```{r Loan Status, echo=FALSE}

# Plot Listings by Loan Status
ggplot(loans, aes(x=loans$LoanStatus)) +
  geom_bar() +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.4)) +
  labs(title='Loan Status',
       x=NULL,
       y='Number of Listings')
```

Most loans appear to be current in their payments or to sucessfully complete. 
It would be interesting to explore correlation between Chargedoff / Defaulted 
loans and ProsperRating or ProsperScore (the Prosper evaluation of a borrower's 
credit worthiness).

### Listing Category
```{r Listing Category, echo=FALSE}
#Plot percent of listings by Category
ggplot(loans, aes(x=loans$ListingCategory, y=..count../n, n=nrow(loans))) +
  geom_bar() +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.4)) +
  labs(title='Listing Category',
       x=NULL,
       y='Percent of Listings')
```

Debt consolidation is the predominant reason on Prosper for borrowing money. 
Over half of all listings are for Debt Consolidation. I would like to explore 
listing category against borrower income to determine if different levels of 
income tend to borrow for different reasons.

### Income Range
```{r Income Range, echo=FALSE}
#Plot percent of listings by income level
ggplot(loans, aes(x=loans$IncomeRange, y=..count../n, n=nrow(loans))) +
  geom_bar() +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x=element_text(angle=90, hjust=1, vjust=0.4)) +
  labs(title='Income Levels of Listings',
       x=NULL,
       y='Percent of Listings')
```

Prosper borrowers appear to be significantly more affluent than the general 
population of the U.S.

| Bracket        |Prosper  |United States  |
|----------------|---------|---------------|
|Less than $25K  |       6%|            35%|
|$25K to $50K    |      28%|            36%|
|$50K to $75K    |      27%|            16%|
|$75K to $100K   |      15%|             6%|
|Over $100K      |      15%|             7%|

(Prosper values exclude borrowers with income not displayed, who are unemployed, 
or who have no income. United States values are for people aged 25-64 with income.)

Prosper borrowers could be older and likely to earn more. Prosper borrowers are 
certainly self-selecting as the type of person who would take a loan out through 
an online lender. Prosper's lending requirements could also bias to more affluent 
borrowers.

Reference: [Personal Income in the United States](https://en.wikipedia.org/wiki/Personal_income_in_the_United_States#Income_distribution)

### Loan Amount
```{r Loan Amount, echo=FALSE}

#boxplot of LoanOriginalamount
n1 <- 
  ggplot(loans, aes(x='', y=LoanOriginalAmount)) +
    geom_boxplot(outlier.shape=NA) +
    geom_jitter(data=dplyr::filter(loans,
      LoanOriginalAmount.outlier.high == TRUE 
      | LoanOriginalAmount.outlier.low == TRUE),
      color='red', width=0.1, alpha=0.1) +
    labs(title='Box Plot of LoanOriginalAmount',
         x=NULL,
         y='LoanOriginalAmount')

#histogram of LoanOriginalamount
n2 <-
  qplot(loans$LoanOriginalAmount,
        geom='histogram',
        binwidth=1000,
        col=I('white'),
        main='Histogram of LoanOriginalAmount',
        xlab='Loan Amount')  

#tabular data of loan amount frequency divisible by 10s, 100s, 1000s, 10000s
n3 <- tibble('Divisor'=integer(),
             'Distinct_Values'=integer(),
             'Number_of_Loans'=integer())

for (i in 1:4) { #powers of 10 math and summary
  loans %>%
  filter(LoanOriginalAmount%%(10^i) == 0) %>%
  summarize(sum=sum(n()), count=n_distinct(LoanOriginalAmount)) %>%
  {add_row(n3, 'Divisor' = (10^i),
                          'Distinct_Values'= .$count,
                          'Number_of_Loans' = .$sum)} -> n3
}


#histogram of last digit of loan amount
n4 <-
  qplot(loans$LoanOriginalAmount%%10,
      geom='histogram',
      binwidth=1,
      col=I('white'),
      main='Histogram of Last Digit',
      xlab='Last Digit')


#place in grid
grid.arrange(n1, n2, tableGrob(n3), n4, ncol=2)


#cleanup
rm(n1, n2, n3, i, n4)
```

#### Loan Amount Summary Statistics
```{r Loan Amount 2, echo=FALSE}
summary(loans$LoanOriginalAmount)
```

```{r Loan Amount 3, echo=FALSE}

loans %>%
  count(LoanOriginalAmount) %>%
  arrange(desc(n)) %>%
  mutate(pct=n/sum(n)) %>%
  head(10)
```

Prosper currently offers loans between $2,000 and $40,000. This data set appears 
to have data before that change was implemented, as loans in the set are bounded 
$1,000-$35,000. There is bias towards "large, round" loan amounts, indicating 
borrowers tend to select an amount near what they need rather than precisely. 
About 74% (84,000 loans of 114,000 total loans) of loans are a multiple of 
thousands.

$4,000 is the most commonly borrowed amount. I'm curious if the amount is also 
associated with a specific type of purchase (other than debt consolidation). 
$15,000 and $10,000 are also frequently occuring amounts. These top three account 
for one-third of all loans, and the top 10 loan amounts account for 61%.

Loan amounts skew right, which is typical of a personal income distribution, and 
I would expect loan amount to have a positive correlation with borrower income.

### Defaults by Cycle
```{r Defaults by Cycle Histogram, echo=FALSE}
ggplot(subset(loans, !is.na(LoanFirstDefaultedCycleNumber))) +
  geom_histogram(aes(x=LoanFirstDefaultedCycleNumber, y=..count..),
                 binwidth=1) +
  labs(title='Number of Loans by First Defaulted Cycle',
       x='First Defaulted Cycle',
       y='Count')
```

*Defaults by Cycle, Deciles*
```{r Defaults by Cycle, Deciles, echo=FALSE}
quantile(loans$LoanFirstDefaultedCycleNumber, probs=seq(0,1,0.1), na.rm=TRUE)
```

Of loans that enter default, 40% do so within 12 billing cycles. 2.8% of defaults 
occur in cycle 1. Cycle 8 is the single largest cycle for defaults, accounting
for 5.7% of defaults. After cycle 8, the number of defaults gradually declines. 

### Median Loan Amount by State
```{r Median Loan Amount by State, echo=FALSE}

#summarize average loan amount by state
avg_loan_by_state <-
loans %>%
  group_by(BorrowerState) %>%
  summarize(n=n(), loan_amount=sum(LoanOriginalAmount)) %>%
  mutate(state_name = tolower(state.name[match(BorrowerState, state.abb)]),
         amount = loan_amount/n)

#load us state map data
us <- map_data('state')

#plot us map with average loan amounts
ggplot() +
  geom_blank(data=us, aes(long, lat)) +
  geom_map(data=us, map=us, aes(map_id=region),
           fill='black', color='black', size=0.01) +
  geom_map(data=avg_loan_by_state, map=us,
           aes(fill=amount, map_id=state_name), 
           color='grey40', size=0.01) +
  coord_map('polyconic') +
  labs(x=NULL, y=NULL, title='Median Loan Amount by State') +
  theme(panel.border = element_blank(),
        panel.background = element_blank(),
        axis.ticks   = element_blank(),
        axis.text    = element_blank()) +
  scale_fill_viridis(name='Amount', option='inferno',
                     label=scales::dollar)

#cleanup
rm(us, avg_loan_by_state)
```

I would expect that average loan amount and median income by state would be
positively correlated. This is the case in some states, such as California,
Texas, and much of the northeast, but does not hold true in many other areas.

For instance, North Dakota has a relatively high median income, but
comparatively low average loan value. Conversely, Louisiana has median income
comparable to other gulf states, but higher average loan values.

This comparison is not rigorous, and I don't expect to dig in to it further in 
this analysis, but it is an interesting map.

Reference: [Census: Median household income, 2012-2016](https://www.census.gov/quickfacts/fact/map/US/INC110216)

# Univariate Analysis
The Prosper data set is well-structured with loan observations on each row and 
many interesting variables in columns. There is enough information to determine 
if data Prosper collects is sufficient to determine credit risk and how
predictive Prosper's model is of repayment.

I performed slight cleanup on the data to:

* convert loan listing category and loan term from numeric to factors.

* add informational boolean values to high and low loan value outliers.

* simplify handling of datetimes.

* simplify column names.

Other than the above, I performed no adjustment to underlying data values.

For each plot above, I provided interesting findings about the individual
variables I explored. I intend to explore a few relationships further:

* relationship between Prosper rating / score of creditworthiness and the \
default rate 

* reason for borrowing money against borrower income level; are reasons \
consistent or varied?

* loan amount vs. income

# Bivariate Plots
### Correlation Matrix
```{r Correlation Matrix, echo=FALSE}
loans %>%
  select(LoanOriginalAmount, Term, BorrowerAPR, EstimatedReturn,
         ProsperRatingNumeric, CreditScoreRangeUpper, RevolvingCreditBalance,
         AvailableBankcardCredit, DebtToIncomeRatio, BankcardUtilization,
         StatedMonthlyIncome, MonthlyLoanPayment, Recommendations,
         Investors, PercentFunded) %>%
  ggcorr(., layout.exp=4, hjust=.95, size=3) +
  labs(title='Correlation Matrix of Selected Variables')
```

Before creating bivariate plots of the variables I found interesting through 
univariate analysis, I decided to create a Pearson correlation matrix to uncover 
potentially interesting relationships that might exist. Some relationships are 
unsurprising, such as Prosper rating being positively correlated with credit
score (higher is better in both cases) and negatively correlated with borrower 
APR (better rated borrowers get lower rates). I will also explore why estimated
returns are negatively correlated with loan amount and Prosper rating.

### Charge Off Rate by Prosper Rating
```{r Charge Off Rate by Prosper Rating, echo=FALSE}
loans %>%
  filter(ListingCreationDate.dt >= june2009 &
           LoanStatus %in% c('Completed', 'Chargedoff')) %>%
ggplot(., aes(x=ProsperRatingAlpha)) +
  geom_bar(aes(fill=LoanStatus), position='fill') +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x=element_text(size=10, angle=30, hjust=1, vjust=1)) +
  scale_fill_brewer(palette='Spectral') +
  labs(title='Charge Off Rate by Prosper Rating',
       x='Prosper Rating',
       y='Percentage',
       fill='Loan Status')
```

To validate my assumption that lower Prosper ratings have higher Charge Off rates,
I plot the percentage of loans that have entered a final state of 'Completed' or 
'Charged Off'. The result is as expected. Risk of charge off steadily increases 
as rating decreases.

### Borrower APR and Estimated Return
```{r Estimated Returns, echo=FALSE}
loans %>%
  filter(ListingCreationDate.dt >= june2009) %>%
ggplot(., aes(x=BorrowerAPR, y=EstimatedReturn)) +
  geom_point() +
  scale_x_continuous(labels=scales::percent) +
  scale_y_continuous(labels=scales::percent) +
  labs(title='Borrower APR relationship to Estimated Return',
       x='Borrower APR',
       y='Estimated Return')
```

Estimated returns are strongly positively correlated to borrower APR. APRs below
20% have almost entirely positive estimated returns, while APRs above 20% have
clusters of positive and negative estimated returns.

The shape of this data leads me to hypothesize that the groupings are influenced 
by another variable. I will further refine this plot in the multivariate section.

### Estimated Returns by Prosper Rating
```{r Percent Return, echo=FALSE}
loans %>%
  filter(ListingCreationDate.dt >= june2009) %>%
  mutate(EstimatedReturnDollars=EstimatedReturn*LoanOriginalAmount) %>%
  group_by(ProsperRatingAlpha) %>%
ggplot(., aes(x=ProsperRatingAlpha, y=EstimatedReturn)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::percent) +  
  theme(axis.text.x=element_text(size=10, angle=30, hjust=1, vjust=1)) +
  labs(title='Estimated Percent Return by Prosper Rating',
       x='Prosper Rating',
       y='Estimated Return')

```

This corroborates the negative correlation between higher (better) Prosper rating 
and estimated percent return. I suspect this will show up in the multivariate 
plot of return, APR, and rating as discussed above. As we look at lower rating 
tranches, median estimated return increases, estimated return shows wider variation,
and among the highest risk ratings, downside risk increases substantially.

```{r Dollar Return, echo=FALSE}
loans %>%
  filter(ListingCreationDate.dt >= june2009) %>%
  mutate(EstimatedReturnDollars=EstimatedReturn*LoanOriginalAmount) %>%
  group_by(ProsperRatingAlpha) %>%
ggplot(., aes(x=ProsperRatingAlpha, y=EstimatedReturnDollars)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::dollar) +  
  theme(axis.text.x=element_text(size=10, angle=30, hjust=1, vjust=1)) +
  labs(title='Estimated Dollar Return by Prosper Rating',
       x='Prosper Rating',
       y='Estimated Return')

```

In dollar terms, the high risk tranche carries the most risk of negative returns. 
B and C ratings have a good balance of highest return and risk. I expect the 
lower dollar return to E and HR ratings are related to the loan amounts they are
able to request.

### Loan Amount by Prosper Rating
```{r Loan Amount by Rating, echo=FALSE}
loans %>%
  filter(ListingCreationDate.dt >= june2009) %>%
  group_by(ProsperRatingAlpha) %>%
ggplot(., aes(x=ProsperRatingAlpha, y=LoanOriginalAmount)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::dollar) +  
  theme(axis.text.x=element_text(size=10, angle=30, hjust=1, vjust=1)) +
  labs(title='Loan Amount by Prosper Rating',
       x='Prosper Rating',
       y='Loan Amount')
```

E and HR rated borrowers have median loan amounts less than half that of AA-C 
rated borrowers. The median loan amount and inter-quartile ranges
of AA-C rated borrowers are all very similar. This makes sense because it is likely 
they are borrowing money for reasons unrelated directly to their credit rating, 
but a better credit rating enables better access to money.

### Loan Amount by Income Range
```{r Loan Amount by Income Range, echo=FALSE}
ggplot(loans, aes(x=IncomeRange, y=LoanOriginalAmount)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::dollar) +  
  theme(axis.text.x=element_text(size=10, angle=30, hjust=1, vjust=1)) +
  labs(title='Loan Amount by Income Range',
       x='Income Range',
       y='Loan Amount')
```

All income ranges took advantage of loans for the lowest amounts Prosper offers, 
but only $100K+ earners borrowed the maximum amounts. Median loan amount increases 
with income.

### Reason for Borrowing by Income
```{r Proportion of Loan Amount by Reason by Income Level, echo=FALSE}

# Filter excluding 'Not displayed' income range,
# percent of spending by category for each individual income range

loans %>%
  filter(!(IncomeRange %in% 'Not displayed')) %>%
  group_by(IncomeRange, ListingCategory) %>%
  summarize(LoanAmountCat=sum(LoanOriginalAmount)) %>%
  mutate(LoanAmountCatPct=LoanAmountCat/sum(LoanAmountCat)) %>%
ggplot(., aes(x=IncomeRange, y=ListingCategory, fill=LoanAmountCatPct)) +
  geom_tile(color='white') +
  coord_fixed(ratio=.5) +
  scale_fill_gradientn(colors=c(scales::muted('blue'), 'red'),
                       label=scales::percent) +
  theme(axis.text.x=element_text(size=10, angle=90, hjust=1, vjust=1)) +
  labs(x='Income Range', y='Listing Category',
       fill='Percent of Dollars',
       title='Loan Dollars for Listings by Income')
```

As shown above, higher earners borrow larger amounts as a whole. Why do they borrow? 
For each income range, the proportion of borrowing for each listing reason was
calculated. The predominant reason for borrowing across almost all income ranges
is Debt Consolidation. Most other reasons are near insignificant for each income range.
Business borrowing is a second area revealed for borrowers with $0 stated income.
This could be to start a business for someone with enough personal wealth to
qualify for the loan.

### Prosper Rating by Income Range
```{r Prosper Rating by Income Range, echo=FALSE}
loans %>%
  filter(ListingCreationDate.dt >= june2009) %>%
ggplot(., aes(x=IncomeRange)) +
  geom_bar(aes(fill=ProsperRatingAlpha), position='fill') +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x=element_text(size=10, angle=30, hjust=1, vjust=1)) +
  scale_fill_viridis(discrete=TRUE, direction=-1) +
  labs(title='Percent of Prosper Ratings by Income Range',
       x='Income Range',
       y='Percentage',
       fill='Prosper Rating')
```

Higher incomes tend to have better ratings. The lowest rating levels quickly
decrease with moderate income and the highest ratings expand. To me, this suggests 
that people generally want to meet their credit obligations as measured by their 
credit rating, and as they're able, they do.

### Loan Amount by Term
```{r Loan Amount by Term, echo=FALSE}
ggplot(loans, aes(x=Term.f, y=LoanOriginalAmount)) +
  geom_boxplot() +
  scale_y_continuous(labels = scales::dollar) +  
  theme(axis.text.x=element_text(size=10, angle=30, hjust=1, vjust=1)) +
  labs(title='Loan Amount by Term',
       x='Loan Term',
       y='Loan Amount')
```

As expected, higher loan amounts are generally borrowed with longer repayment
terms. I think this would be more interesting to explore with additional variables 
to determine if there other influences, so I will include it in multivariate analysis.

# Bivariate Analysis
My bivariate analysis began with a correlation matrix to uncover surpsising
relationships. Most relationships were expected -- variables commonly associated
with credit worthiness (credit score, credit balances available and utilized,
debt to income ratio, income) were well correlated with Prosper's ratings. I
chose to explore how investor returns were negatively correlated with loan
amount and Propser rating.

Estimated return includes interest, collected late fees, and principal loss due 
to default. My own findings corroborate Prosper's self-published
[investor information](https://www.prosper.com/invest): higher ratings yield
lower returns due to lower interest rate. Lower ratings have higher percentage
returns until the high risk group, which has significant prospects of negative
return.

Also of interest was the stability of reason for borrowing. Debt consolidation
was near universal as a predominant reason for taking a loan, and that
percentage of borrowing increased with income.

Next, I plan to explore more about:

* the relationship of borrower APR and estimated return.

* loan amounts by term length.

# Multivariate Plots Section
### Borrowing Behavior
```{r Borrowing Behavior, echo=FALSE}
loans %>%
  filter(!is.na(ProsperRatingAlpha) & ProsperRatingAlpha != '') %>%
ggplot(., aes(x=ProsperRatingAlpha, y=LoanOriginalAmount)) +
  geom_boxplot() +
  scale_y_continuous(label=scales::dollar) +
  coord_flip() +
  facet_grid(. ~ Term.f) +
  theme(axis.text.x=element_text(size=10, angle=30, hjust=1, vjust=1)) +
  labs(title='Prosper Rating and Loan Amount Boxplot by Term Month',
       x='Prosper Rating',
       y='Loan Amount')
```

When compared to the bivariate analysis of Rating and Loan Amount above, we see 
the same general trend of larger loan amount being borrowed for longer terms.
Some additional characteristics of the data emerge:

* High Risk borrowers have only 36-month terms. I'm unaware if this is due to a 
Prosper credit rating policy or due to other general offer terms.

* Borrowers across all risk groups borrowed similar amounts in 12-month terms.

* AA, A, and B groups borrow similar median amounts across all terms.

### Estimated Return and APR by Rating and Loan Amount
```{r Estimated Return and APR by Rating, echo=FALSE}

loans %>%
  filter(ListingCreationDate.dt >= june2009) %>%
ggplot(., aes(x=BorrowerAPR, y=EstimatedReturn)) +
  geom_point(aes(color=ProsperRatingAlpha, size=LoanOriginalAmount), alpha=0.05) +
  guides(color = guide_legend(override.aes = list(alpha = 1))) +
  scale_x_continuous(labels=scales::percent) +
  scale_y_continuous(labels=scales::percent) +
  scale_size_continuous(labels=scales::dollar) +
  labs(title='Estimated Return and APR by Prosper Rating and Loan Amount',
       x='Borrower APR',
       y='Estimated Return',
       color='Prosper Rating',
       size='Loan Amount')
```

As in the bivariate analysis, clear positive correlation between borrower APR
and estimated return remains. Adding Prosper rating shows distinct clustering of
ratings between APRs with little overlap across ratings. This plot also highlights
the potential for negative returns that exists almost exclusively with high risk
borrowers, and the positive yield characteristics of non-prime lending.

# Multivariate Analysis
Adding Rating and loan amount to the estimated return by APR plot highlighted
the interesting relationship of credit-worthiness and risk. A borrower's past
performance on repaying obligations appears to be more indicative of future
performance than what the amount of a given loan is.

The interaction of Prosper rating, loan amount, and term was surprising. I
expected to see loan term length generally increase with borrowed amount,
but it was intersting how consistently similar amounts were borrowed for
similar length of times across several Prosper ratings.


------

# Final Plots and Summary
### Defaults by Cycle
```{r Defaults by Cycle Histogram Final, echo=FALSE}
ggplot(subset(loans, !is.na(LoanFirstDefaultedCycleNumber))) +
  geom_histogram(aes(x=LoanFirstDefaultedCycleNumber, y=..count..),
                 binwidth=1) +
  labs(title='Number of Loans by First Defaulted Cycle',
       x='First Defaulted Cycle',
       y='Count')
```

Risk of default begins to drop after 12 billing cycles.

### Reason for Borrowing by Income
```{r Proportion of Loan Amount by Reason by Income Level Final, echo=FALSE}

# Filter excluding 'Not displayed' income range,
# percent of spending by category for each individual income range

loans %>%
  filter(!(IncomeRange %in% 'Not displayed')) %>%
  group_by(IncomeRange, ListingCategory) %>%
  summarize(LoanAmountCat=sum(LoanOriginalAmount)) %>%
  mutate(LoanAmountCatPct=LoanAmountCat/sum(LoanAmountCat)) %>%
ggplot(., aes(x=IncomeRange, y=ListingCategory, fill=LoanAmountCatPct)) +
  geom_tile(color='white') +
  coord_fixed(ratio=.5) +
  scale_fill_gradientn(colors=c(scales::muted('blue'), 'red'),
                       label=scales::percent) +
  theme(axis.text.x=element_text(size=10, angle=90, hjust=1, vjust=1)) +
  labs(x='Income Range', y='Listing Category',
       fill='Percent of Dollars',
       title='Loan Dollars for Listings by Income')
```

Across income ranges, people borrow money for substantially similar reasons on
Prosper.

### Prosper Rating by Income Range
```{r Prosper Rating by Income Range Final, echo=FALSE}
loans %>%
  filter(ListingCreationDate.dt >= june2009) %>%
ggplot(., aes(x=IncomeRange)) +
  geom_bar(aes(fill=ProsperRatingAlpha), position='fill') +
  scale_y_continuous(labels = scales::percent) +
  theme(axis.text.x=element_text(size=10, angle=30, hjust=1, vjust=1)) +
  scale_fill_viridis(discrete=TRUE, direction=-1) +
  labs(title='Percent of Prosper Ratings by Income Range',
       x='Income Range',
       y='Percentage',
       fill='Prosper Rating')
```

Borrowers with higher incomes tend to have better Prosper ratings.

------

# Reflection
Overall, it was interesting spending so much time learning R, which seems to be
incredibly powerful and complex with many ways to do things that aren't always 
completely intuitive.

Exploring the Prosper data set was interesting. One of the best and worst things
about choosing this data set was the sheer volume of variables that were available.
It might have been easier to focus if there weren't so many choices, or if I knew 
there was a particular feature I was looking for. However, I learned a lot!

In the future, it would be interesting to measure the actual return of completed 
loans against the expected return to see how well Prosper's system helps investors
mitigate risk and maximize returns.
